The yaq Project: \\ Standardized Software Enabling Flexible Instrumentation

### 2023-03-16 SI.tex
si-typos

@@

We believe that Python and the Scientific Python ecosystem is a powerfully simple tool for defining experimental procedures.
Lightweight scripts such as these are encouraged by the \yaq{} project.

@@

@@

@@

While \yaq{} is flexible enough to represent such arrays, it was not specifically designed with large cameras as the primary usecase.
As such, performance does suffer above about 1 Megapixel.
Figure \ref{si:fig:scaling} shows the relationship between number of pixels and time for a \yaq{} message to read the array.
Below about $2^{18}$ (~250,000) pixels, there is virtually no dependence on size, with the standard overhead of a \yaq{} message dominating the time for transfer.
\yaq{} remains usable for up to approximately $2^{20}$ (~1,000,000) pixels.
``Usable'' is a relative term which will depend greatly on context.
Here we generally mean ``seems responsive to a user trying to have feedback on human timescales''.
Applications which require high speed, high throughput cameras are unlikely to be suitable for \yaq{} even with relatively small cameras.
This test was conducted using 32-bit integer arrays.

@@

Finally \yaq{} daemons could be implemented to write arrays locally to disk, rather than transferring over the \yaq{} interface directly.
Anecdotally, many users of Bluesky, EPICS, and related technologies use a similar strategy for large image data.
This has not yet been implemented for \ya{}, and would require clients to have knowledge of how to retrieve and display the images collected, but there is nothing preventing this if throughput is a limitation.
If such behavior became a standard feature desired for cameras, it should be encapsulated in a \yaq{} trait to provide a consistent interface.

>>>>

We believe that Python and the Scientific Python ecosystem is a powerfully simple tool for defining experimental procedures.
Lightweight scripts such as these are encouraged by the \yaq{} project.
The following two sections contain example orchestration scripts for common tasks.

@@

@@

Sometimes you want to acquire data while a peripheral continuously moves, asynchronously with data acquisition.
The following is an example where sample temperature is continuously ramped and data is collected every second.

@@

Imagine that the temperature starts out at room temperature, 25.
We first set the temp client to 100, beginning to heat the sample.
The outer while loop will continue until the temperature reaches our desired end-point of 90.
Each second, the while loop collects the current temperature and the current sensor reading.
This type of experiment is common in the chemical sciences.

Note that we pay no attention to the busy state of the temp client.
When compared with the previous raster scanning example, we see that \yaq{} can be flexible about which hardware operations are kept synchronous and which are allowed to run ``freely'' without blocking other operations.
The ubiquity and simplicity of the busy message allows for this flexibility.

@@

While \yaq{} is flexible enough to represent such arrays, it was not specifically designed with large cameras as the primary usecase.
As such, performance does suffer above about 1 Megapixel.
Figure \ref{si:fig:scaling} shows the relationship between number of pixels and time for a \yaq{} message to read the array.
Below about $2^{18}$ ($\approx250,000$) pixels, there is virtually no dependence on size, with the standard overhead of a \yaq{} message dominating the time for transfer.
\yaq{} remains usable for up to approximately $2^{20}$ ($\approx1,000,000$) pixels.
``Usable'' is a relative term which will depend greatly on context.
Here we generally mean ``seems responsive to a user trying to have feedback on human timescales''.
Applications which require high speed, high throughput cameras are unlikely to be suitable for \yaq{} even with relatively small cameras.
This test was conducted using 32-bit integer arrays.

@@

Finally \yaq{} daemons could be implemented to write arrays locally to disk, rather than transferring over the \yaq{} interface directly.
Anecdotally, many users of Bluesky, EPICS, and related technologies use a similar strategy for large image data.
This has not yet been implemented for \yaq{}, and would require clients to have knowledge of how to retrieve and display the images collected, but there is nothing preventing this if throughput is a limitation.
If such behavior became a standard feature desired for cameras, it should be encapsulated in a \yaq{} trait to provide a consistent interface.
